---
title: "Presentation Ninja"
subtitle: "&#x2694;&#xFE0F; xaringan +<br/>&#x1F60E; xaringanthemer"  
author: 
  - "Yihui Xie"
  - "Garrick Aden-Buie"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.height = 8, fig.width=12)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

class: inverse, center, middle

# Presentation for ETC5523

### Presenter: Hanh Ngo

### Presenter: Yuqi Wang





---
class: center, middle

# Questions we explored

--

### Is there evidence to suggest that the air quality in Melbourne decreased during the summer bush fire season?

--

### How does air quality compare now?

---
class: inverse, center, middle

# Dataset

---
## Source

- Microclimate Sensor Readings

- Microclimate Sensor Locations

## Variables

- PM2.5: Mass density of particles in the air smaller than 2.5 micrometers in diameter.

- PM10: Mass density of particles in the air smaller than 10 micrometers in diameter.

- PM2.5-EPA-24h: Mass density of particles in the air smaller than 2.5 micrometers in diameter, 24 hour rolling average.

- PM10-EPA-24h: Mass density of particles in the air smaller than 10 micrometers in diameter, 24 hour rolling average.

---
class: inverse, center, middle

# Findings

---


### Data cleaning

```{r data, message=FALSE,warning=FALSE,include=FALSE}
library(dplyr)
library(lubridate)
library(naniar)
library(ggplot2)
library(plotly)
library(ggridges)
library(viridis)
library(hrbrthemes)
sensor_location <- read.csv(here::here("data", "Microclimate_Sensor_Locations.csv"))
sensor_reading <- read.csv(here::here("data", "Microclimate_Sensor_Readings.csv"))

```

```{r message=FALSE,warning=FALSE, echo=TRUE, results='hide'}
str(sensor_reading)  # Check structure, notice some factor type

sensor_reading <- sensor_reading %>%  # Change from factor to date time
 mutate(local_time = ymd_hms(local_time),  # Change some factors to character type
         type = as.character(type),
         units = as.character(units))

sensor_reading_location <- sensor_reading %>%  # Merge with sensor location
  left_join(sensor_location, by = "site_id")

pm <- sensor_reading %>%
  filter(type %in%c("PM2.5", "PM10", "PM2.5-EPA-24h", "PM10-EPA-24h")) %>%  # Choose the sensor type for analyzing
  mutate(day = as.Date(local_time,"%Y/%m/%d"))         # mutate day

pm <- pm %>%
  mutate(month = case_when(
    day>= "2019-11-01" & day< "2019-12-01" ~"2019-11",  # mutate month
    day>= "2019-12-01" & day< "2020-01-01" ~"2019-12",
    day>= "2020-01-01" & day< "2020-02-01" ~"2020-01",
    day>= "2020-02-01" & day< "2020-03-01" ~"2020-02",
    day>= "2020-03-01" & day< "2020-04-01" ~"2020-03",
    day>= "2020-04-01" & day< "2020-05-01" ~"2020-04",
    day>= "2020-05-01" & day< "2020-06-01" ~"2020-05",
    day>= "2020-06-01" & day< "2020-07-01" ~"2020-06",
    day>= "2020-07-01" & day< "2020-08-01" ~"2020-07",
    day>= "2020-08-01" & day< "2020-09-01" ~"2020-08",
    day>= "2020-09-01" & day< "2020-10-01" ~"2020-09",
    day>= "2020-10-01" & day< "2020-11-01" ~"2020-10"))
```

---

### Head the tidy data

```{r }
knitr::kable(head(pm,5), format = 'html')
```

---

### The air quality from 2019-11-15 to 2020-10-26

```{r barplot, message=FALSE, warning=FALSE, echo=FALSE}

pm_4 <- pm %>%
  filter(type %in%c("PM2.5", "PM10", "PM2.5-EPA-24h", "PM10-EPA-24h")) %>%
  select("value", "type", "month") %>%
  group_by(month, type) %>%
  mutate(VALUE = mean(value)) 
g4 <- ggplot(pm_4, aes(month, VALUE, fill= type)) +
  geom_bar(stat = "identity", 
position = "dodge") + 
  xlab("Month") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
g4
```


---
### The air quality from 2019-11-15 to 2020-10-26

```{r message=FALSE,warning=FALSE,echo=FALSE}
pm_6 <- pm %>%
  filter(type %in%c("PM2.5-EPA-24h", "PM10-EPA-24h")) %>%
  select("value", "type", "month") %>%
  group_by(month, type)
g5 <-ggplot(pm_6, aes(x=month, y=value, fill=type)) + 
    geom_boxplot() +
    facet_wrap(~type, nrow = 2, scales = "free_y") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
g5
```

---

### Comparing the air quality in January 2020 with October 2020

```{r message=FALSE,warning=FALSE,echo=FALSE}
g6 <- ggplot(pm_6, aes(x=month, y=value, fill=type)) + 
    geom_boxplot() +
    facet_wrap(~type, nrow = 2) +
    ylim(0,100) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
g6
```

---

### Comparing the air quality in January 2020 with October 2020

```{r message=FALSE,warning=FALSE,echo=FALSE}
pm_2 <- pm %>%
  select("value", "type", "day", "local_time") %>%
  filter(type %in%c("PM2.5-EPA-24h", "PM10-EPA-24h")) %>%
  filter(day >= "2019-12-01",
         day < "2020-03-01") %>%
  group_by(day, type) 
g2 <- ggplot(pm_2, aes(local_time, value, color= type)) +
  geom_line()
pm_3 <- pm %>%
  select("value", "type", "day", "local_time") %>%
  filter(type %in%c("PM2.5-EPA-24h", "PM10-EPA-24h")) %>%
  filter(day >= "2020-10-01",
         day < "2020-11-01") %>%
  group_by(day, type) 
g3 <- ggplot(pm_3, aes(local_time, value, color= type)) +
  geom_line()
gridExtra::grid.arrange(g2,g3,nrow=2)
```

---

### Comparing the air quality in January 2020 with October 2020

```{r message=FALSE,warning=FALSE,echo=FALSE}
pm_5 <- pm %>%
  filter(type %in%c("PM10-EPA-24h")) %>%
  select("value", "type", "month") %>%
  group_by(month, type) 

# Plot
ggplot(pm_5, aes(x = value, y = month, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01 ) +
  xlim(15,70) +
  scale_fill_viridis(name = "Mean PM10 [μg/m3]", option = "C") +
  labs(title = 'PM10 in Melbourne') +
  theme_ipsum() 

```

---

class: inverse, center, middle

# Conclusions


---

class: center, middle

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
